Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-07-04 WholePlotTrajectories.tex
IDH graphs

In tropical forests, ecological communities are regularly re-shaped by
natural disturbance events that change both the abiotic environment,
through the fluxes of light, heat and water \citep{Goulamoussene2017},
and the biotic interactions such as competition among species
\citep{Chesson2000, Herault2018}. One of the cornerstone of tropical
forest ecology is to understand the processes and drivers of ecosystems
response to disturbance \citep{White2001, Chazdon2003a}. For now, this
has been largely studied through forest structural parameters such as
aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016} that are rapid and convenient to
measure. These structural parameters have been sucessfully modeled,
giving important insights into the recovery of ecosystem processes and
services \citep{Herault2018}. However the response of forests diversity
and composition is still unclear, albeit it determines the productivity,
stability and functioning of ecosystems \citep{Tilman2014, Liang2016}.
In the short-term, moderate disturbance may lead to positive impacts on
communities diversity, an idea formalized by the intermediate
disturbance hypothesis (IDH) stating a maximized species diversity when
disturbance intensity is not too high
\citep{Molino2001, Kariuki2006a, Berry2008a}.

Validations of the IDH though remain scarce in the long-term and mainly
rely on the analysis of taxonomic richness \citep{Molino2001}. Taxonomic
richness, alone, gives limited or misleading information on forests
recovery and functioning \citep{Martin2015, Chaudhary2016}. More
ecological-meaningful analysis would couple richness to (i) evenness to
reveal the underlying ecological processes and (ii) composition that is
crucial for conservation issues
\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
functional approach accounting for species biological attributes would
directly link communities diversity, composition and redundancy to
ecosystem functioning and to its environmental constraints
\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
respect, the functional trait-based approach that focused on major
traits related to species ecology and thta mediated species performance
in a given environment was sucessfully adopted
\citep{Diaz2005, Villeger2008a}. For instance, the functional approach
revealed in tropical rainforests the deterministic processes entailing,
after disturbance, a functional shift from a dominance of
``conservative'' slow-growing species dealing with scarce resources to
``acquisitive'' fast-growing species with rapid and efficient use of
abundant resources \citep{TerSteege2001, Reich2014, Herault2011}. This
shift is translated into consistent trajectories of key functional
traits related to resource acquisition (leaf and stem traits) and
life-history traits (seed mass, maximum size)
\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Eventually
a complete overview of communities response to disturbance would
encompass the changes in functional redundancy, that quantifies the
amount of shared trait values among species \citep{Carmona2016}. The
high functional redundancy of hyperdiverse tropical forests
\citep{Bellwood2006} mitigates the impacts of species removal on
ecosystem functioning and determines the resilience of communities after
disturbance \citep{Trenbath1999, Elmqvist2003, Diaz2005}.

In this study, we monitored over 30 years the response of 75 ha of
neotropical forest plots set up on a gradient of disturbance intensity,
from 10 to 60\% of ecosystem biomass removed. We made use of a large
functional traits database encompassing major leaf, stem and
life-history traits in order to draw the taxonomic and functional
trajectories in terms of richness, evenness, composition and redundancy
\citep{Lohbeck2015, Guariguata2001}. Specifically, we (i) questioned the
recovery of communities taxonomic and functional characteristics and
identify the underlying assembly processes, (ii) clarified the validity
of the IDH in the long term for tropical forest and elucidated its
declination into different trajectories, and (iii) questioned the length
of complete communities recovery.

@@

Paracou station in French Guiana (5??18'N and 52??53'W) is located in a
lowland tropical rain forest in a tropical wet climate with mean annual
temperature of 26??C, mean annual precipitation averaging 2980
mm.y\textsuperscript{-1} (30-y period) and a 3-month dry season
(\textless{} 100 mm.month\textsuperscript{-1}) from mid-August to
mid-November, and a one-month dry season in March \citep{Wagner2011}.
Elevation ranges between 5 and 50 m and soils correspond to thin
acrisols over a layer of transformed saprolite with low permeability
generating lateral drainage during heavy rains.

@@

The study site corresponds to a tropical rainforest typical of the
Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae,
Lecythidaceae and Sapotaceae. In the twelve experimental plots, all
trees above 10 cm DBH have been mapped and measured annually since 1984.
Trees are first identified with a vernacular name assigned by the forest
worker team, and afterward with a scientific name assigned by botanists
during regular botanical campaigns. In 1984, specific vernacular names
are given to 62 commercial or common species whereas more infrequent
ones were identified under general identifiers only distinguishing trees
and palm trees. From 2003, botanical campaigns have been conducted every
5 to 6 years to identify all trees at the species level but
identification levels still varied among plots and campaigns.

This variability of protocols in time raised methodological issues as
vernacular names usually correspond to different botanical species. It
resulted in significant taxonomic uncertainties that had to be
propagated to composition and diversity metrics. The uncertainty
propagation was done through a Bayesian framework reconstituting
complete inventories at genus level from real incomplete ones on the
basis of vernacular/botanical names association. Vernacular names were
replaced through multinomial trials based on the association probability
\(\big[\alpha_1, \alpha_2,???, \alpha_3\big]\) observed across all
inventories between each vernacular name \emph{v} and the species
\(\big[s_1, s_2, ???, s_N\big]\):

\begin{align}
M_v\Big(\big[s_1, s_2, ???, s_N\big],\big[\alpha_1, \alpha_2,???, \alpha_3\big]\Big) \nonumber
\end{align}

@@

Six functional traits representing leaf economics (leaves thickness,
toughness, total chlorophyll content and specific leaf area) and wood
economics (wood specific gravity and bark thickness), and life-history
traits (maximum specific height and seed mass) came from the BRIDGE
project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were
assessed from a selection of individuals located in nine permanent plots
in French Guiana, including two in Paracou, and comprised 294 species
pertaining to 157 genera. Missing trait values were filled using
multivariate imputation by chained equation \citep{Mice2011}.
Imputations were restricted within genus, or family when samples were
too scarce, in order to account for the phylogenetic signal. Whenever a
species was not in the dataset, it was attributed a set of traits values
randomly sampled among species of the next higher taxonomic level (same
genus or family). As seed mass information was classified into classes,
no data filling process was applied and analyses were restricted to the
414 botanical species recorded.

All composition and diversity metrics were obtained after 50 iterations
of the taxonomic uncertainty propagation framework.

@@

To counter taxonomic uncertainties due to the variability of botanical
identification levels (in space) and protocols (in time), the taxonomic
composition and diversity analysis were conducted at the genus level.
Taxonomic and functional trajectories of community composition were
followed in a two-dimensional NMDS ordination space. Two NMDS were
conducted to map either taxonomic or functional composition, the later
based on the 7 leaf, stem and life history traits (without seed mass
classes). NMDS were performed using abundance-based (Bray-Curtis)
dissimilarity measures. Trajectories along time were reported through
the euclidean distance between the target inventories and the reference
inventories in 1989, \emph{i.e.} 2-3 years after disturbance. Univariate
trajectories of the leaf, stem and life-history traits were also
visualized with the community weighted means (CWM)
\citep{Diaz2007, Garnier2004}. Species seed mass corresponded to 5
classes of increasing mass, seed mass trajectories were therefore
reported as the proportion of each class in the inventories (Supp Mat
XX).

The taxonomic diversity was reported through species richness and the
evenness, \emph{i.e} the Hill number translation of the Simpson index
\citep{Hill1973}. These two indices belong to the set of HCDT or
generalized entropy, respectively corresponding to the 0 and 2 order of
diversity (q), recomended for diversity studies
\citep{Patil1982, Tothmeresz1995, Marcon2015}. The functional diversity
was reported using the Rao index of quadratic entropy which combines
species abundance distribution and average pairwise dissimilarity based
on all functional traits.

The impacts of initial disturbance were tested with the spearman rank
correlation between the extrema of taxonomic and functional metrics
reached over the 30 years and the initial \%AGB removed. The were
besides analysed through the linear correlations between Simpson and Rao
diversities and the initial \%AGB removed at 10, 20 and 30 years after
disturbance.

The functional redundancy was measured as the overlap among species in
community functional space \citep{Carmona2016}. The samples of the trait
database were first mapped in a 2-dimensional plan with a PCA analysis.
Then, multivariate kernel density estimator associated with individual
trees returned species traits probability distribution (TDP). Species
TDP weighted by species abundance were eventually summed for each
community: the functional redundancy was the sum of TDPs overlap,
expressed as the average number of species that could be removed from
without reducing the functional space (see appendix I for a more
comprehensive sheme).

@@

The impact of disturbance intensity was examined through the linear
correlation between (i) the \%AGB removed and (ii) taxonomic richness,
evenness and functional diversity after 10, 20 and 30 years
\ref{fig:IDHplot}. The correlation with disturbance intensity was weak
for XXXXXXXXXX, for the taxonomic evenness (\(R^2<0.25\)) while highly
significant for functional diversity (\(0.60<R^2<0.75\)) whatever the
time since disturbance.

@@

Communities taxonomic and functional composition appeared resilient,
following similar hump-shaped trajectories starting to return towards
pre-disturbance composition after 30 years. The taxonomic differences
among communities, marked before disturbance by the distinct starting
points on the NMDS axis 2, were maintained throughout recovery
trajectories. More than commonly thought, post-disturbance trajectories
depended on community initial composition, that partly determined the
pool of recruited species and that constrained the trajectories towards
the initial composition. The high resilience of communities taxonomy
revealed that species not belonging to the pre-disturbance community
were hardly recruited because of the commonness of dispersal limitation
among tropical tree species (Svenning, Wright JECOL 2005).

>>>>

In tropical forests, ecological communities are regularly re-shaped by
natural disturbance events changing both the abiotic environment,
through the fluxes of light, heat and water \citep{Goulamoussene2017},
and the biotic interactions such as competition among species
\citep{Chesson2000, Herault2018}. One of the cornerstone of tropical
forest ecology is to understand the processes and drivers of ecosystems
response to disturbance \citep{White2001, Chazdon2003a}. For now, this
has been largely studied through forest structural parameters such as
aboveground biomass, tree height or stem density
\citep{Piponiot2016, Rutishauser2016} that are rapid and convenient to
measure. These structural parameters have been sucessfully modeled,
giving important insights into the recovery of ecosystem processes and
services \citep{Herault2018}. However the response of forests diversity
and composition is still unclear, albeit it determines the productivity,
stability and functioning of ecosystems \citep{Tilman2014, Liang2016}.
In the short-term, moderate disturbance may lead to positive impacts on
communities diversity, an idea formalized by the intermediate
disturbance hypothesis (IDH) stating a maximized species diversity when
disturbance intensity is not too high
\citep{Molino2001, Kariuki2006a, Berry2008a}.

Validations of the IDH though remain scarce in the long-term and mainly
rely on the analysis of taxonomic richness \citep{Molino2001}. Taxonomic
richness, alone, gives limited or misleading information on forests
recovery and functioning \citep{Martin2015, Chaudhary2016}. More
ecological-meaningful analysis would couple richness with (i) evenness
that would reveal underlying ecological processes and (ii) composition
that is crucial for conservation issues
\citep{Magurran1988, Lavorel2002, Bellwood2006}. Furthermore, a
functional approach accounting for species biological attributes would
directly link communities diversity, composition and redundancy to
ecosystem functioning and to its environmental constraints
\citep{Violle2007b, Moretti2009, Baraloto2012a, Scheiter2013}. In that
respect, the functional trait-based approach that focus on major traits
related to species ecology and mediate species performance in a given
environment was sucessfully adopted \citep{Diaz2005, Villeger2008a}. For
instance, the functional approach revealed in tropical rainforests the
deterministic processes entailing, after disturbance, a functional shift
from a dominance of ``conservative'' slow-growing species dealing with
scarce resources to ``acquisitive'' fast-growing species with rapid and
efficient use of abundant resources
\citep{TerSteege2001, Reich2014, Herault2011}. This shift is translated
into the trajectories of key functional traits related to resource
acquisition (leaf and stem traits) and life-history traits (seed mass,
maximum size)
\citep{Wright2004, TerSteege2006, Westoby2006a, Chave2009b}. Eventually
a complete overview of communities response to disturbance would
encompass the changes in functional redundancy, that quantifies the
amount of shared trait values among species \citep{Carmona2016}. The
high functional redundancy of hyperdiverse tropical forests
\citep{Bellwood2006} mitigates the impacts of species removal on
ecosystem functioning and determines the resilience of communities after
disturbance \citep{Trenbath1999, Elmqvist2003, Diaz2005}.

In this study, we monitored over 30 years the response of 75 ha of
neotropical forest plots set up on a gradient of disturbance intensity,
from 10 to 60\% of ecosystem above-ground biomass (AGB) removed. We made
use of a large functional traits database encompassing major leaf, stem
and life-history traits in order to draw the taxonomic and functional
trajectories in terms of richness, evenness, composition and redundancy
\citep{Lohbeck2015, Guariguata2001}. Specifically, we (i) questioned the
recovery of communities taxonomic and functional characteristics and
identified the underlying assembly processes, (ii) clarified the
validity of the IDH in the long term for tropical forest and elucidated
its translation into different trajectories, and (iii) questioned
community recovery time.

@@

Paracou station in French Guiana (5\textdegree 18'N and
52\textdegree 53'W) is located in a lowland tropical rain forest in a
tropical wet climate with mean annual temperature of 26\textdegree C,
mean annual precipitation averaging 2980 mm.y\textsuperscript{-1} (30-y
period) and a 3-month dry season (\textless{} 100
mm.month\textsuperscript{-1}) from mid-August to mid-November, and a
one-month dry season in March \citep{Wagner2011}. Elevation ranges
between 5 and 50 m and soils correspond to thin acrisols over a layer of
transformed saprolite with low permeability generating lateral drainage
during heavy rains.

@@

The study site corresponds to a tropical rainforest typical of the
Guiana Shield with a dominance of Fabaceae, Chrysobalanaceae,
Lecythidaceae and Sapotaceae. In the twelve experimental plots, all
trees above 10 cm DBH have been mapped and measured annually since 1984.
Trees are first identified with a vernacular name assigned by the forest
worker team, and afterward with a scientific name assigned by botanists
during regular botanical campaigns. In 1984, specific vernacular names
were given to 62 commercial or common species whereas more infrequent
ones were identified under general identifiers only distinguishing trees
and palm trees. From 2003, botanical campaigns have been conducted every
5 to 6 years to identify all trees at the species level but
identification levels still varied among plots and campaigns.

This variability of protocols in time raised methodological issues as
vernacular names usually correspond to different botanical species. It
resulted in significant taxonomic uncertainties that had to be
propagated to composition and diversity metrics. The uncertainty
propagation was done through a Bayesian framework reconstituting
complete inventories at genus level from real incomplete ones on the
basis of vernacular/botanical names association. Vernacular names were
replaced through multinomial trials based on the association probability
\(\big[\alpha_1, \alpha_2,..., \alpha_3\big]\) observed across all
inventories between each vernacular name \emph{v} and all species
\(\big[s_1, s_2,..., s_N\big]\):

\begin{align}
M_v\Big(\big[s_1, s_2,..., s_N\big],\big[\alpha_1, \alpha_2,..., \alpha_3\big]\Big) \nonumber
\end{align}

@@

Six functional traits representing leaf economics (leaves thickness,
toughness, total chlorophyll content and specific leaf area) and wood
economics (wood specific gravity and bark thickness), and life-history
traits (maximum specific height and seed mass) came from the BRIDGE
project \footnote{http://www.ecofog.gf/Bridge/}. Trait values were
assessed from a selection of individuals located in nine permanent plots
in French Guiana, including two in Paracou, and comprised 294 species
pertaining to 157 genera. Missing trait values (10\%) were filled using
multivariate imputation by chained equation \citep{Mice2011}.
Imputations were restricted within genus or family when samples were too
scarce, in order to account for the phylogenetic signal. Whenever a
species was not in the dataset, it was attributed a set of trait values
randomly sampled among species of the next higher taxonomic level (same
genus or family). As seed mass information was classified into classes,
no data filling process was applied and analyses were restricted to the
414 botanical species recorded.

All composition and diversity metrics were obtained after 50 iterations
of the uncertainty propagation framework.

@@

To counter taxonomic uncertainties due to the variability of botanical
identification levels (in space) and protocols (in time), the taxonomic
composition and diversity analysis were conducted at the genus level.
Taxonomic and functional trajectories of community composition were
followed in a two-dimensional NMDS ordination space. Two NMDS using
abundance-based (Bray-Curtis) dissimilarity measures were conducted to
map either taxonomic or functional composition, the later based on the 7
leaf, stem and life history traits (without seed mass classes).
Trajectories along time were reported through the euclidean distance
between the target inventories and the reference inventories in 1989,
\emph{i.e.} 2 years after disturbance. Univariate trajectories of the
leaf, stem and life-history traits were also visualized with the
community weighted means (CWM) \citep{Diaz2007, Garnier2004}. Species
seed mass corresponded to 5 classes of increasing mass, seed mass
trajectories were therefore reported as the proportion of each class in
the inventories (Supp Mat XX).

The taxonomic diversity was reported through species richness and
evenness, \emph{i.e} the Hill number translation of the Simpson index
\citep{Hill1973}. These indices belong to the set of HCDT or generalized
entropy, respectively corresponding to the 0 and 2 order of diversity
(q), recomended for diversity studies
\citep{Patil1982, Tothmeresz1995, Marcon2015}. The functional diversity
was reported using the Rao index of quadratic entropy which combines
species abundance distribution and average pairwise dissimilarity based
on species functional traits.

The impacts of initial disturbance were tested with the spearman rank
correlation between the extrema of taxonomic and functional metrics
reached over the 30 years and the initial \%AGB removed. The were
besides analysed through the linear correlations between (i) taxonomic
and functional richness, evenness and diversity and (ii) the initial
\%AGB removed at 10, 20 and 30 years after disturbance.

The functional redundancy was measured as the overlap among species in
community functional space \citep{Carmona2016}. The samples of the trait
database were first mapped in a 2-dimensional plan with a PCA analysis.
Then, multivariate kernel density estimator associated with individual
trees returned species traits probability distribution (TDP). Species
TDP weighted by species abundance were eventually summed for each
community. Community functional redundancy was the sum of TDPs overlap,
expressed as the average number of species that could be removed from
without reducing the functional space (see appendix I for a more
comprehensive sheme).

@@

The impact of disturbance intensity was examined through the linear
correlation between (i) the \%AGB removed and (ii) taxonomic richness,
evenness and functional diversity after 10, 20 and 30 years
\ref{fig:IDHplot}. The correlation with disturbance intensity was weak
for te taxonomic richness (\(-0.09<R^2<-0.06\)), for the taxonomic
evenness (\(R^2<0.25\)) while highly significant for functional
diversity (\(0.60<R^2<0.75\)) whatever the time since disturbance.

@@

Communities taxonomic and functional composition appeared resilient,
following similar hump-shaped trajectories starting to return towards
pre-disturbance composition after 30 years. The taxonomic differences
among communities, marked before disturbance by the distinct starting
points on the NMDS axis 2, were maintained throughout recovery
trajectories. More than commonly thought, post-disturbance trajectories
depended on community initial composition, that partly determined the
pool of recruited species and constrained the trajectories towards the
initial composition. The high resilience of communities taxonomy
revealed that species not belonging to the pre-disturbance community
were hardly recruited because of the commonness of dispersal limitation
among tropical tree species (Svenning, Wright JECOL 2005).

### 2018-07-04 withBubs.Rmd
IDH graphs

## Community diversity: richness and evenness

@@

```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
load("DB/ReplacementTraj_ForGraphs")

TaxoTraj(CompleteTaxo)

@@

```{r FunDiv, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="Trajectories of the Rao functional diversity over 30 years after disturbance. Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
load("DB/FunctionalTraj_ForGraphs")

plotDiv(CompleteFun,remove=TRUE)
mtext("Rao diversity",side=3,adj=0,line=1)
mtext("Years since disturbance",side=1,line=2.2,adj=1)
legend("right",inset=c(-0.28,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")
```

@@

```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1/1.5,fig.cap="xxx to be changed xxx Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance. Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals. Lower panels, Relationship between the initial \\%AGB removed and Simpson (ref:Boldc) and Rao (ref:Boldd) diversities 10, 20 and 30 years after disturbance.", fig.env="figure*"}
load("DB/SimpsonDraw");load("DB/FunctionalTraj_ForGraphs")
load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")

par(mfrow=c(1,2),mar=c(3,3,2,1),oma=c(2,1,2,3),no.readonly = T)
plotDiv(SimpsonTaxo)
mtext("Simpson diversity",side=3,adj=0,line=1)
plotDiv(CompleteFun,remove=TRUE)
mtext("Rao diversity",side=3,adj=0,line=1)
mtext("Years since disturbance",side=1,line=2.2,adj=1)
legend("right",inset=c(-0.28,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")

plotIDH(Data=Sim,AgbLoss=AGBloss)
plotIDH(Data=Rao,AgbLoss=AGBloss)
mtext("initial AGB lost (%)",side=1,line=2.2,adj=1)
legend("right",inset=c(-0.28,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")

@@

## Community Composition

@@

```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="Trajectories of functional redundancy within the initial functional space over 30 years after disturbance. Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
load("DB/Redundancy_restricted")

@@

1. IDH in Space (Disturbance gradient)
The IDH consistently (predicted) the response of functional diversity to disturbance, but the theory was disproved regarding communities taxonomic richness and evenness.
In contrast the taxonomic richness and evenness were decoupled from the disturbance intensity, questioning the consistency of the IDH as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998].

2. IDH in Time
The humped-shaped trajectories of functional diversity along time 
Still, the taxonomic trajectories were blurried by the hampered recovery of infrequent species that entailed after disturbance a decrease in the taxonomic richness and an increase of the evenness [@Hubbell2001;@Chave2004].

## The functional redundancy, key of the taxonomic resilience

The recovery of the most infrequent species followed the lottery recruitment rules [@Busing2002] and was hampered by the increasing competition among species, implying long term recovery [@Trenbath1999;@Elmqvist2003;@Diaz2005].

Following disturbance the space and resources made available corresponded to a rapid decrease of the functional redundancy within the initial functional space.
The high light, space and nutrient niches were rapidly filled by species, mainly the most dominant and frequent ones.
Thereafter competitive exclusion emerged following the filling of the niches and limited the recruitment of infrequent species [@Busing2002].
The recovery of communities taxonomic richness and evenness relied upon the recovery of the initial functional redudancy.
Although underway 30 years after disturbance the recovery remained unachieved and after 30 years the recovery it was underway but remained unachieved.
This alteration of the functional redundancy meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific communities, with less species and more pioneers [@Haddad2008; @Burslem2000; @Martin2013].
Besides the slowed recovery of rare species increased the risks to loose cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
Apart from the functional characteristics considered here, infrequent species mighthave unique functions i nthe ecosystem or be a key for some fauna.

@@

Our study revealed communities cyclic recovery after disturbance allowing the resilience of their functioning and taxonomic composition with the maintenance of initial differences among communities.
Communities functional evenness was enhanced for  20 years after disturbance through the enrichment of the communities with pioneers and light-demanding species,in accordance with the IDH.
The IDH, though, poorly predicted the disturbance impact on communities taxonomic richness and evenness that were blurried by the emergence of competitive exclusion along time.
The resilience of tropical forests proved consistent although spread over several decades. 
Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Gourlet-Fleury2005].
As the trajectories highlighted the recruitment processes proved central for communities response to disturbance and closer focus on demographical drivers of communities response would clarify the fate of the future forests.

Communities functional characteristics are the most direct link between biodiversity and ecosystem functioning and their resilience here meant the recovery of ecosystem processes and services on a cycle of several decades [@Diaz2005;@Guariguata2001].

>>>>

## Communities richness and evenness

@@

```{r DivTaxo, out.width = '100%', fig.asp = 1.5/3, echo=FALSE, fig.cap="Trajectories of the difference to the 1989 inventories (5 years after disturbance) over 30 years after disturbance of plots communities (ref:Bolda) Richness, (ref:Boldb) Shannon and (ref:Boldc) Simpson diversities. Trajectories correspond to the median (solid line) and 0.025 and 0.975 percentile (gray envelope) observed after 50 iteration of the taxonomic uncertainty propagation. Initial treatments are represented by solid lines colors with green for control, blue for T1,orange for T2 and red for T3.", fig.env="figure*"}
load("DB/ReplacementTraj_ForGraphs")
load("DB/FunctionalTraj_ForGraphs")

par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
TaxoTraj(CompleteTaxo)
plotDiv(CompleteFun,remove=TRUE)
mtext("(c) Rao diversity",side=3,adj=0,line=1)
mtext("Years since disturbance",side=1,line=2.2,adj=1)
legend("right",inset=c(-0.34,0),xpd=NA,legend=c("T0","T1","T2","T3"),col=ColorsTr,lwd=2.5,bty="n",title="Treatment")

@@

@@

```{r IDHplot, out.width = '100%', echo=FALSE, fig.asp = 1.5/3,fig.cap="Upper panels, Trajectories of the taxonomic evenness (Simpson diversity) (ref:Bolda) and Rao functional diversity (ref:Boldb) over 30 years after disturbance. Colors are treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals. Lower panels, Relationship between the initial \\%AGB removed and Simpson (ref:Boldc) and Rao (ref:Boldd) diversities 10, 20 and 30 years after disturbance.", fig.env="figure*"}
load("DB/RichnessIDH");load("DB/SimpsonIDH");load("DB/RaoIDH");load("DB/LostAGB")

par(mfrow=c(1,3),mar=c(5,2,4,2),oma=c(1,1.5,1,1),no.readonly=TRUE)
plotIDH(Data=Rich,AgbLoss=AGBloss)
mtext("(a) Taxonomic Richness",line=1,side=3,adj=0,cex=0.8)
plotIDH(Data=Sim,AgbLoss=AGBloss)
mtext("(b) Taxonomic Evenness",line=1,side=3,adj=0,cex=0.8)
plotIDH(Data=Rao,AgbLoss=AGBloss)
mtext("(c) Functional diversity",line=1,side=3,adj=0,cex=0.8)
mtext("initial AGB lost (%)",side=1,line=2.2,adj=1)
legend("right",inset=c(-0.34,0),xpd=NA,legend=c("10","20","30"),col=colyear,lwd=2.5,bty="n",title="Years")

@@

## Communities Composition

@@

```{r RedFun_rest, out.width = '100%', echo=FALSE, fig.cap="TTrajectories of functional redundancy within the initial functional space over 30 years after disturbance. Colors are disturbance treatments: green (control), blue (T1), orange (T2), red (T3) with shaded areas the credibility intervals.", fig.env="figure"}
load("DB/Redundancy_restricted")

@@

The IDH consistently predicted the response of taxonomic richness to disturbance. The IDH translated in first an increase of richness with diturbance intensity and then a decrease above an intensity threshold (20-25% of initial AGB lost). The taxonomic evenness was somewhat decoupled from disturbance intensity, as already observed in the Guiana Shield [@Baraloto2012a] and in Bornean tropical forests [@Cannon1998]. In the Guiana Shield, the pool of true pioneer species that recruited after disturbance is restricted to a few common general (e.g. Cecropia spp., Vismia spp.) and the post-disturbance richness and evenness, in the taxonomic space, will only depend on the relative proportion of these recruits and the surviving species. -> explain what happens at low, intermediate, high intensity. 
-> functionally diversity: pioneer specis are functionally highly different, we still do not how diverse they are. Their increase in abundance is enough to increase the overall FDiv.

2. IDH in Time
a. first the drop and then gain. quite similar drop, highly different species gain.
b. short-lived pioneers versus long-lived pioneers dynamics linked to habitat diversity
c. porbably, long-lived come after short-lived for high disturbance while they come together for intermediate disturbance.
d. link to fuctional diversity, the decoupling in time between high and low disturbance intensity.

## The functional redundancy, key of the system resilience

1. drop in functional redundancy due to random species loss and replacement by pioneer species ( high light, space and nutrient niches) that do not belong to the initial functional space. the highest the disturbance, the highest the loss and replacement, the deepest the drop in FR. - first 15 years
2. 2nd step: the system started losing pioneer species (less competitive) and replaced them by not-pioneer (either long-lived or end-succesional sp, more competitive), FR increased.
3. end of the process: all functionally different species are lost and replaced by individuals belonging to species from the pre-dicturbance community.
-> long-process. hard to see how long it may last. The recovery of the most infrequent species followed the lottery recruitment rules [@Busing2002] mplying long term recovery [@Trenbath1999;@Elmqvist2003;@Diaz2005]. Although underway 30 years after disturbance the recovery remained unachieved and after 30 years the recovery it was underway but remained unachieved.

This alteration of the functional redundancy likely meant a lower resilience of pre-disturbance communities and higher chances to see the persistence of disturbance-specific communities, with less functionzl redundancy and more pioneers [@Haddad2008; @Burslem2000; @Martin2013].
Besides the slowed recovery of rare species increased the risks to loose cornerstone specie, with unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].
Apart from the functional characteristics considered here, infrequent species might have unique functions i nthe ecosystem or be a key ressource or habitat for some fauna.

@@

[Our study revealed communities cyclic recovery after disturbance allowing the resilience of their functioning and taxonomic composition with the maintenance of initial differences among communities.
Communities functional evenness was enhanced for  20 years after disturbance through the enrichment of the communities with pioneers and light-demanding species,in accordance with the IDH.
The IDH, though, poorly predicted the disturbance impact on communities taxonomic richness and evenness that were blurried by the emergence of competitive exclusion along time.]
The resilience of tropical forests proved consistent although spread over several decades. 
Still, the disturbance impact on communities redundancy cautioned against the risks of infrequent species loss and the persistence of disturbance-specific communities [@Gourlet-Fleury2005]. and we know that regularly disturbed communities are less efficient to recover after disturbance that undisturbed ones..
